% Abstract: AI4MATH 2026 workshop submission.
\begin{abstract}
Lean 4 with Mathlib ships strong general automation (\texttt{aesop},
\texttt{simp}, \texttt{linarith}, \texttt{polyrith}, \texttt{nlinarith},
\texttt{positivity}, \texttt{measurability}, \texttt{bound},
\texttt{gcongr}). It does not ship a hammer for applied
mathematics. The moment a goal mentions a martingale, a stopping
time, a production function, an ordinary differential equation
(ODE) Lyapunov candidate, a Wasserstein distance, or a quantum
density operator, the generic tactics have nothing to apply.
We present \texttt{pythia}, a Lean 4 tactic library for
applied-mathematics proof automation. The headline tactic
\texttt{pythia!} orchestrates a nine-rung deterministic ladder.
The rungs are \texttt{stat\_simp}, the
\texttt{linarith}/\texttt{nlinarith}/\texttt{polyrith} chain,
\texttt{positivity}, \texttt{aesop} on the registered
\texttt{Pythia} ruleset, the \texttt{@[stat\_lemma]} cascade,
\texttt{z3\_check}, \texttt{cvc5\_check}, \texttt{vampire\_check}
or \texttt{e\_check}, and \texttt{disprove}. Each rung tries
fail-fast with a bounded budget. The verbose variant
\texttt{pythia?} reports the closing rung and per-rung wall-clock
timing for diagnostic use. The library ships ten registered
tactics, attribute-driven registries (\texttt{@[stat\_lemma]},
\texttt{@[stats\_ineq]}, \texttt{@[prob\_simp]},
\texttt{@[stat\_simp]}, \texttt{@[anytime\_valid\_lemma]}, plus
the per-domain calculator typeclass), and a fifteen-domain proof
library with thirty-two cross-domain theorems carrying matched
Python sim runners (property-based testing, deterministic sweeps,
mutation testing). Domains span probability and statistics,
biology, chemistry, economics, control theory, optimal transport,
quantum information, stochastic calculus, game theory,
thermodynamics, numerical analysis, queueing, and operations
research. Every public theorem is axiom-clean against the standard
Lean kernel set ($\texttt{propext}$, $\texttt{Classical.choice}$,
$\texttt{Quot.sound}$). The cross-prover ladder is offline-first
and free of large language models (LLMs): \texttt{z3\_check},
\texttt{cvc5\_check}, \texttt{vampire\_check}, and \texttt{e\_check}
dispatch to local Satisfiability Modulo Theories (SMT) and
Automated Theorem Prover (ATP) binaries with kernel-checked
reconstruction in the CoqHammer style of
\citet{czajkakaliszyk2018coqhammer}. LLM-augmented closure lives
in a separate companion software development kit (SDK) that the
library does not depend on. Hard structural Mathlib gaps are
optionally routed to the Aristotle automated theorem prover
\citep{harmonic2025aristotle} for development-time restocking,
with audit gates against vacuous-truth closures. The library is
pinned to Lean 4.28 + Mathlib v4.28.0 and released open-source.
\end{abstract}
