\appendix

\section{Library inventory}
\label{sec:appendix-inventory}

The shipped library at the AI4MATH 2026 submission cut spans the
modules below. Each line lists module path and one-line summary.
All listed modules are axiom-clean against
$\mathcal{A}_{\text{core}} = \{\texttt{propext},
\texttt{Classical.choice}, \texttt{Quot.sound}\}$ unless explicitly
flagged as scaffold.

\subsection{Anytime-valid inference}
\begin{itemize}[leftmargin=1.3em,itemsep=0.2em]
\item \texttt{Pythia.HowardRamdasCS}: Howard-Ramdas Ville bound for
  self-normalised confidence sequences (rate
  $\eta_{\mathrm{HR}}(b) = \sqrt{b \log 2}$). Axiom-clean.
\item \texttt{Pythia.BettingCS}: betting-CS coverage certificate
  (rate $\eta_{\mathrm{betting}}(b) = 1/\sqrt{b\log 2 + 1}$).
  Axiom-clean.
\item \texttt{Pythia.BettingStrategy}: betting-strategy admissibility
  scaffolding for the \texttt{anytime\_valid} dispatch.
\item \texttt{Pythia.VilleSupermartingale}: countable-time and
  finite-horizon Ville inequalities. The
  \texttt{ville\_supermartingale}, \texttt{ville\_ineq}, and
  \texttt{ville\_supermartingale\_finite} lemmas are the workhorse.
\item \texttt{Pythia.EDetector}: e-process formalism, with
  \texttt{combine\_eprocesses\_avg} (average-merge) as the workhorse
  closer. Anytime-valid Type-I error control via Markov-at-stopping.
\end{itemize}

\subsection{Concentration}
\begin{itemize}[leftmargin=1.3em,itemsep=0.2em]
\item \texttt{Pythia.SubGaussianMG}: sub-Gaussian martingale
  concentration. The Chernoff exponent $\exp(-\tau^2/(2\sigma^2 N))$
  bound. Axiom-clean.
\item \texttt{Pythia.SubGamma}: sub-gamma tail bound. Axiom-clean.
\item \texttt{Pythia.MGFBoundedSubGamma}: bounded-RV
  Bernstein-Bennett MGF bound,
  $\mathbb{E}[\exp(\lambda X)]
   \leq \exp(\sigma^2 \lambda^2 / (2(1 - b\lambda/3)))$
  for centred $|X| \leq b$. Axiom-clean.
\item \texttt{Pythia.Bernstein}: iid and martingale Bernstein
  applications.
\end{itemize}

\subsection{Matrix concentration}
\begin{itemize}[leftmargin=1.3em,itemsep=0.2em]
\item \texttt{Pythia.MatrixLieb}: Lieb concavity step
  \citep{lieb1973} and operator-monotone log composition. Scaffold.
  Dependent operator-trace inequalities route to Aristotle.
\item \texttt{Pythia.MatrixBernstein}: iid matrix-Bernstein head
  theorem \citep{tropp2012usermatrix}. Scaffold.
\item \texttt{Pythia.MatrixBernsteinFull}: martingale extension.
  Scaffold.
\end{itemize}

\subsection{Other modules}
\begin{itemize}[leftmargin=1.3em,itemsep=0.2em]
\item The \texttt{Pythia.Control} subpackage carries
  \texttt{LyapunovDiscrete}, the discrete-time Lyapunov stability
  result for linear systems.
\item \texttt{Pythia.Queueing.ErlangB}: Erlang-B blocking formula.
\item \texttt{Pythia.Queueing.LittlesLaw}: Little's law steady-state
  identity.
\item \texttt{Pythia.TimeSeries.} \texttt{WoldDecomposition} \citep{wold1938}.
\item \texttt{Pythia.Risk.CVaR} \citep{rockafellaruryasev2000}: the
  Rockafellar-Uryasev variational characterisation of CVaR.
\item \texttt{Pythia.Asymptotics.DeltaMethod},
  \texttt{Pythia.Asymptotics.} \texttt{DeltaMethodMulti}: scalar and
  multivariate delta method.
\item \texttt{Pythia.InfoTheory.} \texttt{BretagnolleHuberBinary}: binary-case
  total-variation bound via the Bhattacharyya bridge.
\item The \texttt{Pythia.MeasureTheory} subpackage carries
  \texttt{OptionalStoppingUnbounded} and \texttt{PathMeasureRN}.
  Both are structural support for the Ville chain.
\end{itemize}

\section{Worked examples}
\label{sec:appendix-worked}

Each example below is a literal regression test from the
\texttt{pythia} test suite. Every \texttt{example} compiles
axiom-clean against $\mathcal{A}_{\text{core}}$.

\subsection{anytime\_valid: Ville-bound closer}
\label{sec:appendix-worked-av}

The countable-time variant closes the canonical Ville maximal
inequality on a non-negative supermartingale.

\begin{lstlisting}
import Pythia.Tactic.AnytimeValid
open MeasureTheory ProbabilityTheory ENNReal Pythia

example {Omega : Type*} {m0 : MeasurableSpace Omega}
    {mu : Measure Omega} [IsFiniteMeasure mu]
    {f : Nat -> Omega -> Real}
    {F : Filtration Nat m0}
    (hsup : Supermartingale f F mu)
    (hnn  : forall t omega, 0 <= f t omega)
    (hint : Integrable (f 0) mu)
    {c : Real} (hc : 0 < c) :
    mu {omega | exists t : Nat, f t omega >= c}
      <= (Integral.integral mu (f 0)).toNNReal / c.toNNReal := by
  anytime_valid
\end{lstlisting}

The tactic dispatches via the
\texttt{ville\_supermartingale} \texttt{exact} path. Hypotheses are
discharged by \texttt{assumption}. No registered rule is consulted.
The kernel-checked term reduces under
$\{\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}\}$.

\subsection{z3\_check: linear-real-arithmetic with reconstruction}
\label{sec:appendix-worked-z3}

A three-step linear chain over $\mathbb{R}$. Z3 is queried as a
ranking filter. \texttt{linarith} produces the actual proof term.

\begin{lstlisting}
import Pythia.Tactic.Z3Check
open Pythia

example {a b c : Real} (h1 : a < b) (h2 : b <= c) :
    a < c := by
  z3_check
\end{lstlisting}

Trace: the encoder produces the SMT-LIB query
\texttt{(set-logic QF\_LRA) ... (assert (< v\_a v\_b)) (assert
(<= v\_b v\_c)) (assert (not (< v\_a v\_c))) (check-sat)}, which Z3
returns \texttt{unsat}. \texttt{linarith} then constructs the
Farkas certificate inside Lean. Z3 cannot smuggle axioms. If
\texttt{linarith} disagreed, the tactic would fail loud.

\subsection{pythia: chain-rule fall-through to aesop}
\label{sec:appendix-worked-pythia-fall}

A goal with no matching \texttt{@[stat\_lemma]} but a closed-form
chain in \texttt{linarith}. Stage 1 (aesop with the \texttt{Pythia}
ruleset) declines. Stage 2 (the cleanup chain combining
\texttt{simp}, \texttt{omega}, \texttt{linarith}, and
\texttt{positivity}) closes.

\begin{lstlisting}
import Pythia.Tactic.Pythia
open Pythia

example (a b c d : Real)
    (h1 : a <= b) (h2 : b <= c) (h3 : c <= d) :
    a <= d := by
  pythia
\end{lstlisting}

The dispatcher's \texttt{first}-combinator tries aesop, observes no
applicable \texttt{@[stat\_lemma]} closer, then tries the cleanup
chain. \texttt{linarith} closes the chain in one step.

\subsection{Registry-extend: tagging a user theorem}
\label{sec:appendix-worked-extend}

User extensions consist of tagging library lemmas with the
appropriate attribute. Below: a trivial real-arithmetic identity
registered for \texttt{pythia} via \texttt{@[stat\_lemma]}.

\begin{lstlisting}
import Pythia.Tactic.Pythia
open Pythia

@[stat_lemma]
theorem add_zero_real (x : Real) : x + 0 = x := by ring

-- pythia now closes goals matching the registered lemma:
example (x y : Real) : (x + 0) + (y + 0) = x + y := by
  pythia
\end{lstlisting}

The attribute re-elaborates as
\texttt{@[aesop safe apply (rule\_sets := [Pythia])]}. The lemma
then joins the \texttt{Pythia} aesop ruleset and is consulted by
\texttt{pythia} on every subsequent invocation. No fork or rebuild
of the tactic implementation is required.

\section{Aristotle vs DSPv2 vs Opus 4.6 (wider subset)}
\label{sec:appendix-comparison}

Table~\ref{tab:wide-comparison} reports a wider subset of FormalAVS
\citep{anonymous2026formalavs} per-tier closure rates for the three
solver families. Aristotle is one-attempt-per-target. DSPv2 and Opus
are pass@5 at $N{=}5$ on the $48$-target slate (T0 through T2 plus T3).
Cells are \emph{(closures)/(targets)} at the tier denominator, with
T4 (library-gap) and T5 (capability-ceiling) appended.

\begin{table}[h]
\centering
\small
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccccc}
\toprule
\textbf{Solver} & T0 & T1 & T2 & T3 & T4 & T5 \\
\midrule
Aristotle              & 7/7 & 33/34 & 7/8 & 6/6 & 0/4 & 0/1 \\
DSPv2-7B GPTQ-int8     & 3/7 & 8/34  & 0/8 & 1/6 & 0/4 & 0/1 \\
Opus 4.6               & 0/7 & 13/34 & 0/8 & 0/6 & 0/4 & 0/1 \\
\bottomrule
\end{tabular}%
}
\caption{Wider per-tier comparison from FormalAVS. Aristotle: one
attempt per target. DSPv2 and Opus: pass@5 at $N{=}5$. T4 is
solver-hit-Mathlib-gap (every solver $0/4$ for the same library-gap
reason). T5 is the all-solver capability-ceiling target. The 6/6 vs
1/6 vs 0/6 column at T3 (cross-function inequality wall) is the
benchmark's cleanest solver-specific signal. The single DSPv2 T3
closure is the $\eta\textsc{vector}{=}\sqrt{2}\,\eta\textsc{hr}$
identity (the \emph{etavector} target). Kimi K2.5 and Mistral
Large 3 also close this one in the wider drafter sweep. Source: dataset paper
\citep{anonymous2026formalavs}, Tables 1-2 reproduced verbatim.}
\label{tab:wide-comparison}
\end{table}

\paragraph{Per-target capability-wall tail.}
Table~\ref{tab:wide-tail} reports the per-target closure matrix on
the five capability-wall tail problems used in the dataset paper to
classify solver regimes.

\begin{table*}[t]
\centering
\small
\begin{tabular}{lcccc}
\toprule
\textbf{Target} & Aristotle & Kimi K2.5 & Sonnet 4.6 & Gemini 3 Pro \\
\midrule
\texttt{betting\_is\_kelly\_optimal}\textsuperscript{$\star$} & $1/1$ & $1/1$ & $1/1$ & $1/1$ \\
\texttt{c\_vector\_eq\_sqrt\_two\_mul\_c\_HR}     & $1/1$ & $1/1$ & $0/5$ & $1/2$ \\
\texttt{c\_sharp\_ranking}                        & $1/1$ & $0/9$ & $0/5$ & $1/3$ \\
\texttt{quantized\_entropy\_bound}                & $1/1$ & $0/9$ & $0/5$ & $0/3$ \\
\texttt{equivalence\_break\_at} \texttt{\_finite\_precision} & $0/3$\textsuperscript{$\dagger$} & $0/19$ & $0/15$ & $0/13$ \\
\bottomrule
\end{tabular}
\caption{Per-target capability-wall tail. Cells are
\emph{(closures)/(attempts)}. Five distinct solver regimes appear:
universal closure, majority closure, drafter-monopoly (Gemini's
\texttt{gcongr}-prior closes \texttt{c\_sharp\_ranking}),
Aristotle-monopoly, and solver-unreached. Source:
\citet{anonymous2026formalavs}, Table 4.
\textsuperscript{$\star$}Companion-archive entry, not in the
$48$-slate drafter sweep. \textsuperscript{$\dagger$}Aristotle: two
sessions timed out at the 20-minute compute gate. One returned a
counterexample on the as-stated form (missing symmetry hypothesis,
\citep{anonymous2026formalavs} \S Aristotle-refute).}
\label{tab:wide-tail}
\end{table*}

\section{Reproduction}
\label{sec:appendix-repro}

\paragraph{Toolchain pin.}
Lean 4.28 + Mathlib v4.28.0. The repository's
\texttt{lean-toolchain} file declares the exact toolchain. The
Mathlib dependency in \texttt{lakefile.toml} pins the v4.28.0 release
tag.

\paragraph{Build.}
The library URL is redacted for double-blind review (full URL in
camera-ready). Reproduction commands once the URL is supplied:
\begin{verbatim}
git clone <pythia-url-redacted>
cd pythia
lake exe cache get
lake build
\end{verbatim}

\paragraph{Axiom audit.}
Each shipped module ends with a \texttt{\#print axioms} block. The
expected audit set is
$\{\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}\}$.
Any deviation fails the audit.

\paragraph{FormalAVS benchmark.}
The full benchmark is reported in
\citet{anonymous2026formalavs}. Reproduction commands for the
benchmark are in that paper's reproduction appendix. The benchmark
URL is redacted for double-blind review.

\section{Related work, extended}
\label{sec:appendix-related}

\subsection{Lean 4 tactic infrastructure}
Mathlib ships a tactic suite the \texttt{pythia} dispatcher composes
with. \texttt{linarith} produces Farkas certificates and
\texttt{polyrith} routes polynomial-ideal goals via Sage with
kernel-checked reconstruction. \texttt{nlinarith} extends
\texttt{linarith} to a small non-linear fragment. Sign goals route
through \texttt{positivity} and $\sigma$-algebra-membership goals
through \texttt{measurability}. Generalised-congruence monotonicity
is discharged by \texttt{gcongr}. Monotone-composition goals route
through \texttt{bound}, which uses an attribute-driven rule
database.
The closest Mathlib relative to \texttt{pythia} is \texttt{bound}:
the \texttt{@[stats\_ineq]} attribute literally re-tags lemmas as
\texttt{@[bound]}, so the underlying \texttt{bound} dispatch picks
them up automatically. Aesop \citep{aesop2023} supplies the
\texttt{rule\_sets} mechanism the \texttt{pythia} dispatcher uses.
\texttt{pythia} is therefore a thin domain-registration layer over
\texttt{aesop} plus a Z3 oracle.

\subsection{Sledgehammer (Isabelle)}
Sledgehammer
\citep{paulsonblanchette2010threeyears, blanchettesledgehammer2011}
dispatches Isabelle/HOL goals to Vampire, E, Z3, CVC4, with proof
reconstruction inside Isabelle's kernel via Metis or
\texttt{smt}-tactic replay. The reconstruction discipline is the
template for both CoqHammer and \texttt{z3\_check}. There is no Lean
4 equivalent at the time of writing. \texttt{pythia} does not
attempt the Isabelle Sledgehammer's full premise-selection front
end. The library ships only the linear-real-arithmetic instance via
\texttt{z3\_check}.

\subsection{CoqHammer (Coq)}
CoqHammer \citep{czajkakaliszyk2018coqhammer} dispatches Coq goals
to Vampire, E, Z3, CVC4 with kernel-checked reconstruction. Premise
selection in CoqHammer is k-NN over a featurised representation of
the Coq context. The reconstruction step is a small certifying tactic
(\texttt{sauto}) that replays the ATP's witness inside the Coq
kernel. \texttt{z3\_check}'s reconstruction via \texttt{linarith}
mirrors this template at small scale. The external prover supplies
the verdict and the in-kernel tactic supplies the proof term. The
kernel is the final arbiter. We do not claim a full Lean-side reimplementation of
CoqHammer's premise-selection layer.

\subsection{Statistical formalisation libraries}
Mathlib's \texttt{Probability.Martingale} module supplies Doob
decomposition and Azuma's inequality but does not instantiate the
canonical confidence-sequence families, the per-family rate
functions, or Ville's inequality for the betting construction.
The \texttt{pythia} library closes that gap. The FormalAVS
benchmark \citep{anonymous2026formalavs} measures the result.
Coq's mathcomp.analysis and the Isabelle/HOL probability
formalisations cover overlapping but disjoint sub-areas, and
are not attribute-driven hammers.

