\section{Drafter swarm and the audit pipeline}
\label{sec:aristotle}

Many of the theorems in the library involve structural gaps that
Mathlib does not currently close, and many require a proof shape
that the general tactics cannot synthesise on their own. We
seeded the corpus with help from a swarm of contemporary
drafters working in parallel: the Aristotle automated theorem
prover \citep{harmonic2025aristotle}, the DeepSeek-Prover-V2 7B
model \citep{deepseekprover2024v2}, Claude Sonnet 4.6, and Claude
Opus 4.6. Drafters propose candidate proofs. The Lean kernel
accepts or rejects them. None of the drafters is an oracle. The
only authority in the loop is the Lean kernel and the axiom
audit. We use Z3 as a deterministic SMT decision procedure
inside the headline tactic ladder
(Section~\ref{sec:tactic-surface}), not as a drafter, and treat
its verdict as a verifiable witness rather than as a trusted
result.

\subsection{Per-drafter contribution}
\label{sec:aristotle-contribution}

Different drafters close different proof shapes well. Aristotle
is strongest on dominated-convergence side conditions and
operator-theoretic intermediate lemmas. DeepSeek-Prover-V2 is
strongest on local arithmetic identities once the tactic chain
has reduced the goal to a closed-form algebraic obligation.
Sonnet 4.6 is strongest on tactic-orchestration proofs that
need the right Mathlib lemma name plus a short
\texttt{ring}/\texttt{linarith} chain. Opus 4.6 is strongest on
multi-step proofs that need a non-trivial intermediate
construction. No single drafter dominates across the fifteen
domains. The library is built from the union of their successful
contributions, kernel-checked.

\subsection{Submission pattern}
\label{sec:aristotle-pattern}

Each drafter call follows the same packaging contract. State the
theorem in Lean with a \texttt{sorry} placeholder. Package the
project without the Lean build cache (\texttt{.lake} excluded).
Submit to the drafter (Aristotle is a tarball-submission API,
DeepSeek-Prover-V2 runs on a local GPU, Sonnet and Opus run via
chat-completion endpoints). Monitor for completion. Integrate
the returned proof. Run the axiom audit and the vacuous-truth
refusal pass before merging.

\subsection{Vacuous-truth refusal}
\label{sec:aristotle-vacuous}

A drafter can close a theorem by deriving a contradiction from
its hypotheses and concluding anything. The Lean kernel accepts
such a proof: it is sound. The theorem statement is the bug.

\paragraph{Case (2026-04-26).}
We submitted a candidate \texttt{wald\_wolfowitz\_optimal}
theorem. Aristotle returned a closure that combined a Bool
partition with a log-boundary constraint to force both
$\alpha \geq 1$ and $\alpha < 1$. The kernel accepted the proof.
The audit flagged the hypothesis chain as self-contradictory.
We rejected the closure, restated the theorem in two-measure
form (separating the null and alternative measures explicitly),
and resubmitted. The two-measure form was the intended target.
The original Bool-partition form was a statement bug.

\paragraph{Audit policy.}
The library treats vacuous closures as theorem-statement bugs,
not as prover successes. Every drafter-returned proof is audited
for contradictory premises before integration. Closures flagged
as vacuous block the corresponding theorem from the axiom-clean
library until the statement is reformed. The policy is not
specific to any drafter. The same audit applies to candidate
proofs from Aristotle, DeepSeek-Prover-V2, Sonnet, Opus, and
human contributions that close via an unnoticed hypothesis
contradiction.

\subsection{Z3 as a deterministic decision procedure}
\label{sec:aristotle-z3}

The \texttt{z3\_check} tactic in the headline ladder dispatches
linear-real-arithmetic goals to Z3 \citep{demoura2008z3}. Z3 is
not a drafter and is not part of the swarm. Z3 is a
deterministic decision procedure that returns a verdict (unsat
of the negated goal). The verdict is reconstructed inside Lean
by \texttt{linarith}. The Lean kernel is the final word. The
pattern is the CoqHammer
\citep{czajkakaliszyk2018coqhammer} reconstruction strategy
adapted to Lean 4. Z3 cannot smuggle axioms. \texttt{linarith}
is axiom-clean and the reconstruction step certifies the
witness inside the kernel. If \texttt{linarith} cannot
reconstruct, the \texttt{z3\_check} call fails loud rather than
silently trusting Z3. The same reconstruction pattern applies
to \texttt{cvc5\_check}, \texttt{vampire\_check}, and
\texttt{e\_check}.

\subsection{When to invoke a drafter}
\label{sec:aristotle-when}

The library convention is to attempt local closure first. If the
goal closes by \texttt{rfl}, \texttt{simp}, \texttt{positivity},
\texttt{linarith}, \texttt{nlinarith}, \texttt{polyrith}, the
\texttt{stat\_lemma} cascade, or one of the
deterministic-decision-procedure rungs in \texttt{pythia!}, no
drafter call is made. Drafters are reserved for hard structural
Mathlib gaps: measure-theoretic $\sigma$-additivity steps,
dominated-convergence side conditions, matrix-concentration
intermediate lemmas, and similar obligations for which a
hand-Mathlib closure either does not exist or would require an
order of magnitude more dispatch. The vacuous-closure audit is
the gate on integration. The dispatch threshold is only the
call decision.
