\section{Comparison and discussion}
\label{sec:discussion}

\subsection{Comparison to existing automation}
\label{sec:discussion-comparison}

\paragraph{aesop.}
Aesop \citep{aesop2023} is the workhorse forward-search tactic in
Lean 4. \texttt{pythia} declares two named aesop rule sets
(\texttt{Pythia} and \texttt{Pythia.AnytimeValid}) and uses aesop in
stage 1 of the dispatcher chain. \texttt{pythia} is not a replacement
for aesop. It is a statistics-domain registry plus a Z3 oracle plus a
vacuous-closure audit, layered on top of aesop.

\paragraph{Sledgehammer.}
Sledgehammer \citep{paulsonblanchette2010threeyears,
blanchettesledgehammer2011} dispatches Isabelle/HOL goals to external
ATPs (Vampire, E, Z3, CVC4) and reconstructs the proof. There is no
direct Lean 4 equivalent. \texttt{pythia} addresses the
statistics-domain case via \texttt{z3\_check} for
linear-real-arithmetic and Aristotle as a structural oracle. A
general-purpose Lean Sledgehammer is out of scope for this paper.

\paragraph{CoqHammer.}
CoqHammer \citep{czajkakaliszyk2018coqhammer} dispatches Coq goals
to external ATPs with kernel-checked reconstruction. \texttt{z3\_check}
in \texttt{pythia} adopts the same reconstruction template
(witness from external prover, in-kernel checking via
\texttt{linarith}). CoqHammer is the design template. We do not
claim a Lean-side reimplementation of CoqHammer's full premise
selection.

\paragraph{Mathlib's existing tactic suite.}
Mathlib supplies \texttt{linarith}, \texttt{polyrith},
\texttt{nlinarith}, \texttt{positivity}, \texttt{measurability},
\texttt{bound}, \texttt{gcongr}. None of these is a statistics-domain
hammer. \texttt{stats\_ineq} is the closest \texttt{pythia} relative
of the Mathlib \texttt{bound} family: the
\texttt{@[stats\_ineq]} attribute re-tags lemmas as
\texttt{@[bound]}, so \texttt{bound}'s mature dispatch picks them up
automatically. \texttt{pythia} composes with all of the listed
Mathlib tactics. The fall-through chain at stage 2 of the dispatcher
combines \texttt{simp}, \texttt{omega}, \texttt{linarith}, and
\texttt{positivity}. The chain is the explicit composition.

\subsection{Aristotle vs DSPv2 vs Opus 4.6 (illustrative)}
\label{sec:discussion-comparison-empirical}

Table~\ref{tab:illustrative-comparison} reports four FormalAVS
problems \citep{anonymous2026formalavs} on which Aristotle, DSPv2
(DSPv2-7B GPTQ-int8, a quantised local-GPU prover), and Opus 4.6 (an
API generalist drafter) were each given a single-attempt protocol
(Aristotle is one session per problem. DSPv2 and Opus run pass@5
at $N=5$ on the $48$-slate.) The full benchmark numbers are in the
dataset paper. The
point of this subset is to expose the four solver regimes that the
benchmark tail surfaces. Per-problem solve outcomes are
kernel-checked.

\begin{table}[h]
\centering
\small
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lccc}
\toprule
\textbf{Problem} & \textbf{Aristotle} & \textbf{DSPv2} & \textbf{Opus 4.6} \\
\midrule
\texttt{betting\_is\_kelly\_optimal}\textsuperscript{$\star$}      & $1/1$ & n/a    & n/a    \\
\texttt{etavector\_eq\_sqrt\_two\_mul\_etahr}      & $1/1$ & $1/5$  & $0/5$ \\
\texttt{c\_sharp\_ranking}                         & $1/1$ & $0/5$  & $0/5$ \\
\texttt{equivalence\_break\_at} \texttt{\_finite\_precision} & $0/3$\textsuperscript{$\dagger$} & $0/5$ & $0/5$ \\
\bottomrule
\end{tabular}%
}
\caption{Illustrative four-problem subset of FormalAVS. Each cell is
\emph{(closures)/(attempts)}. Aristotle is one session per problem.
DSPv2 and Opus are pass@5 at $N{=}5$. The four rows expose four
distinct solver regimes: universal closure (top), majority closure
(\texttt{etavector}, the $\sqrt{2}$ identity Kimi and Mistral also
close at $N{=}5$), Aristotle-monopoly (\texttt{c\_sharp\_ranking}),
and solver-unreached frontier (\texttt{equivalence\_break}).
\textsuperscript{$\star$}Companion-archive entry, not in the
$48$-slate drafter sweep, included for the universal regime.
\textsuperscript{$\dagger$}Aristotle returned a counterexample on the
as-stated form (missing symmetry hypothesis). The corrected form
remains open. Full per-tier benchmark numbers in
\citet{anonymous2026formalavs}.}
\label{tab:illustrative-comparison}
\end{table}

The wider table on the full benchmark subset is in
Appendix~\ref{sec:appendix-comparison}.

\subsection{Limitations}
\label{sec:discussion-limitations}

Three concrete gaps. (i) Domain scope: \texttt{pythia} is
statistics-domain only. It does not address general proof automation
in Lean 4. (ii) Toolchain pin: the library ships pinned to Lean
4.28 + Mathlib v4.28.0. Mathlib evolves quickly. Each major Mathlib
release typically requires a porting pass. (iii) Aristotle
dependency: the hardest closures route to a proprietary external
prover. The \texttt{z3\_check} oracle is open-source. The Aristotle
oracle is not. The library is usable without Aristotle for any goal
that closes via the \texttt{pythia} dispatcher chain or
\texttt{z3\_check}. A subset of structural Mathlib gaps requires
either Aristotle or a hand proof.
