\section{Introduction}
\label{sec:intro}

Working scientists rely on the Python scientific stack to do
applied mathematics. SciPy, NumPy, statsmodels, and PyMC give a
probabilist domain-aware functions to manipulate distributions,
estimators, and martingales. A control theorist composes
\texttt{control.lqr} with NumPy's linear-algebra routines to
analyse stability. A computational biologist runs Hardy-Weinberg
equilibrium under the same pytest setup as the rest of their
codebase. The empirical workflow is mature and the test
discipline is sharp. Property-based testing exercises the
empirical statement under random inputs in the modelled regime.
Deterministic parameter sweeps cover a grid. Mutation testing
flips one operator or one constant in the empirical model and
confirms the test suite catches the change.

Lean 4 with Mathlib has grown rapidly on the formal-proof side
\citep{moura2021lean, mathlib2020community, mathlib2025}. A
broad general tactic suite ships with it. Forward search lives in
\texttt{aesop} and rewriting in \texttt{simp}. Arithmetic is
split across \texttt{linarith}, \texttt{polyrith}, and the
small non-linear fragment in \texttt{nlinarith}. Sign goals
route through \texttt{positivity} and the $\sigma$-algebra
fragment through \texttt{measurability}. Monotone composition
uses \texttt{bound}. Generalised congruence uses \texttt{gcongr}. Each tactic is general-purpose. None of
them know about each other, and none of them know about the
domains a working scientist actually proves things in. A
biologist proving Hardy-Weinberg has to compose \texttt{ring}
with a hand-picked Mathlib lemma. An economist proving constant
returns to scale for Cobb-Douglas has to chain real-power-cast
rewrites with \texttt{ring}. A queueing theorist proving
Little's law has to assemble the steady-state identity by hand. The orchestration is the user's
job, and the user has to be a Lean expert before they can be a
domain expert.

The other side of the disconnect is the test discipline. A
working scientist's empirical specification of Hardy-Weinberg
lives as a Python function with a hypothesis test and a
parameter sweep, exercised under pytest in CI. A formal Lean
statement of the same theorem typically lives in a separate
file, in a separate language, with no shared tests. The two
specifications drift. No library closes the gap on both sides
at once. There is no domain-aware tactic surface paired with a
Python test layer that keeps the empirical and formal
statements aligned.

\paragraph{Contribution.}
We present \texttt{pythia}, a Lean 4 library that delivers a
domain-aware proof toolkit and an empirical test layer in one
package. The library is organised around three contributions.

(1) \emph{A corpus of thirty-two cross-domain theorems across
fifteen applied-math fields.} Probability and statistics covers
anytime-valid inference, sub-Gaussian and sub-gamma
concentration, and matrix concentration. Biology covers
Hardy-Weinberg equilibrium, Lotka-Volterra, SIR conservation,
Michaelis-Menten saturation, and mass-action kinetics. Chemistry
covers Arrhenius rate-law positivity and Henderson-Hasselbalch.
Economics covers Cobb-Douglas constant returns to scale, CRRA
concavity, CAPM linear pricing, and Walras' law. Control theory
covers scalar and discrete-time Lyapunov stability. Optimal
transport covers discrete $L_1$ Wasserstein non-negativity.
Quantum information covers qubit von Neumann entropy
non-negativity. Stochastic calculus covers the It\^o isometry,
the Burkholder, Davis, and Gundy inequality at $p{=}2$, and
Robbins-Siegmund and Robbins-Monro convergence. Game theory
covers the von Neumann two-strategy minimax inequality.
Thermodynamics covers the Carnot upper bound. Numerical
analysis covers Newton iteration positivity,
Picard-Lindel\"of, KKT conditions, and Kahan summation. Queueing
covers Erlang-B and Little's law. Operations research covers
the OR specialisation of Little's law. Every public theorem is
axiom-clean against
$\{\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}\}$.

(2) \emph{Matched Python sim runners on every theorem.}
Property-based testing exercises the empirical statement under
random inputs in the modelled regime. Deterministic sweeps
exercise it on a parameter grid. Mutation testing flips one
operator or one constant in the empirical model and confirms the
test suite catches the change. The Lean statement and the Python
empirical model share test names and regime-of-validity
declarations so the two specifications stay aligned.

(3) \emph{An eleven-tactic surface plus six registries.} The
headline tactic \texttt{pythia!} orchestrates a nine-rung
deterministic ladder. The first five rungs are Mathlib-side:
\texttt{stat\_simp} on the curated probability and ENNReal simp
set, the \texttt{linarith}/\texttt{nlinarith}/\texttt{polyrith}
chain, \texttt{positivity}, \texttt{aesop} on the registered
\texttt{Pythia} ruleset, and the \texttt{@[stat\_lemma]}
shape-dispatch cascade. The next four are external decision
procedures with kernel-checked reconstruction. Z3 and CVC5
handle Satisfiability Modulo Theories. Vampire and E handle
first-order logic. The final rung \texttt{disprove} catches
vacuous-truth goal statements via a Z3 model query on the
negation. Each rung tries fail-fast with a bounded budget, and
the first rung to close wins. The verbose variant
\texttt{pythia?} reports the closing rung and per-rung
wall-clock timing. The cross-prover ladder is offline-first and
free of large language models. Six registries
(\texttt{@[stat\_lemma]}, \texttt{@[stat\_simp]},
\texttt{@[stats\_ineq]}, \texttt{@[prob\_simp]},
\texttt{@[anytime\_valid\_lemma]}, plus the per-domain
\texttt{DomainCalculator} typeclass) let users register new
lemmas. The headline tactic dispatches them automatically.

\paragraph{How the corpus was built.}
Many of the domain theorems in the library involve structural
gaps that Mathlib does not currently close, and many require a
proof shape that the general tactics cannot synthesise on their
own. We seeded the corpus with help from a swarm of contemporary
drafters: the Aristotle automated theorem prover
\citep{harmonic2025aristotle}, the DeepSeek-Prover-V2 7B model
\citep{deepseekprover2024v2}, Claude Sonnet 4.6, and Claude Opus
4.6. Each candidate proof produced by a drafter is submitted to
the Lean kernel for checking. Drafters are not oracles. The only
authority in the loop is the Lean kernel and the axiom audit.
Each candidate proof goes through three filter stages before it
enters the library. The kernel checks the proof. The axiom
auditor confirms the proof reduces under the standard kernel
set. A vacuous-truth refusal pass catches proofs that closed by
deriving a contradiction from a malformed theorem statement.
Different drafters contribute different theorems, with no single
drafter dominating across the fifteen domains.

\paragraph{Paper organisation.}
Section~\ref{sec:tactic-surface} describes the ten-tactic surface
and the registry attributes.
Section~\ref{sec:library} groups the theorem corpus and the
matched Python sim runners by domain.
Section~\ref{sec:aristotle} describes how candidate proofs from
the drafter swarm are filtered into the library and the
vacuous-truth refusal policy that drove our restatement of
\texttt{wald\_wolfowitz\_optimal}.
Section~\ref{sec:discussion} compares against \texttt{aesop},
Sledgehammer, CoqHammer, and Mathlib's existing tactic suite,
and states limitations.

\paragraph{Reproduction and code.}
The \texttt{pythia} library is open-source. The repository URL
is redacted for double-blind review. Axiom audit transcripts
and reproduction recipes are in
Appendix~\ref{sec:appendix-repro}.
