\section{Library content}
\label{sec:library}

The shipped \texttt{pythia} library spans a tactic layer plus a
proof-content layer of fifteen applied-math domains. The full
manifest at the time of writing carries thirty-two cross-domain
theorems with matched Python sim runners (property-based testing,
parameter sweeps, mutation testing) plus four hundred and sixty
total theorem and lemma declarations across the
\texttt{Pythia/} source tree, fifty-five of which are tagged
\texttt{@[stat\_lemma]} and registered into the headline cascade.
Every theorem listed here is axiom-clean against
$\{\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}\}$
under the Lean 4.28 + Mathlib v4.28.0 toolchain pin. We summarise
by group. The appendix lists modules in detail.

\subsection{Anytime-valid inference}
\label{sec:library-av}

Howard-Ramdas confidence sequences \citep{howard2021time}, betting-CS
\citep{waudbysmith2024estimating}, and the vector-valued extension
\citep{whitehouse2025timeuniform} are each backed by a Ville-bound
specialization in the \texttt{anytime\_valid} tactic. The
\texttt{Pythia.HowardRamdasCS} and \texttt{Pythia.BettingCS} modules
expose the admissibility certificates of the e-process formalism of
\citet{ramdasgrunwaldvovkshafer2023}. The supporting Ville machinery
lives in \texttt{Pythia.VilleSupermartingale} (countable-time and
finite-horizon variants).

\subsection{Concentration and tail bounds}
\label{sec:library-conc}

\texttt{stats\_ineq} dispatches sub-Gaussian and sub-gamma tail
bounds, the Bernstein-Bennett MGF bridge, and Hoeffding-style
inequalities for bounded random variables. Modules:
\texttt{Pythia.SubGaussianMG} (sub-Gaussian martingale concentration),
\texttt{Pythia.SubGamma} (sub-gamma tail bound),
\texttt{Pythia.MGFBoundedSubGamma} (the bounded-RV Bernstein-Bennett
MGF bound, $\mathbb{E}[\exp(\lambda X)] \leq \exp(\sigma^2 \lambda^2 /
(2(1 - b\lambda/3)))$ for centred $|X| \leq b$), and
\texttt{Pythia.Bernstein} (the iid and martingale Bernstein
applications). The MGF bridge composes with \texttt{prob\_simp} for
the integral side and with general Mathlib infrastructure
(\texttt{measurability}, \texttt{integrability}) for side conditions.

\subsection{E-detectors and merging}
\label{sec:library-edet}

The e-process formalism of
\citet{ramdasgrunwaldvovkshafer2023, shinramdasrinaldo2024}
yields uniform-time validity by combining e-processes via averaging.
\texttt{Pythia.EDetector} declares \texttt{EProcess} as a non-negative
process bounded in expectation under the null and \texttt{EDetector}
as the first-passage stopping rule with anytime-valid Type-I error
$\leq \alpha$. The \texttt{combine\_eprocesses\_avg} lemma (average
merge) is the workhorse. Product-merge variants register alongside.
The average-merge case is axiom-clean.

\subsection{Matrix concentration}
\label{sec:library-matrix}

Matrix Bernstein \citep{tropp2012usermatrix} and the Lieb concavity
inequality \citep{lieb1973} are the two matrix-side modules.
\texttt{Pythia.MatrixLieb} states the Lieb concavity step and the
operator-monotone log composition.
\texttt{Pythia.MatrixBernstein} carries the iid matrix-Bernstein head
theorem. \texttt{Pythia.MatrixBernsteinFull} carries the
matrix-martingale extension. The dependent operator-trace inequalities
route through Aristotle as oracle (Section~\ref{sec:aristotle}).

\subsection{Cross-domain coverage (15 fields)}
\label{sec:library-other}

The cross-domain manifest covers fifteen fields with Lean proofs
and matched Python sim runners. Each entry is a closed-form
theorem with property-based-test mutation coverage in the
companion runner.

\begin{itemize}[leftmargin=1.3em,itemsep=0.2em]
\item \textbf{Probability and statistics:} the anytime-valid /
  concentration / matrix sections above plus \texttt{Pythia.Risk.}
  \texttt{CVaR}~\citep{rockafellaruryasev2000} and
  \texttt{Pythia.Asymptotics.} \texttt{DeltaMethod} (scalar) /
  \texttt{DeltaMethodMulti} (multivariate).
\item \textbf{Information theory:} \texttt{Pythia.InfoTheory.}
  \texttt{BretagnolleHuberBinary} (binary-alphabet bound),
  \texttt{Pythia.InfoTheory.DataProcessing} (data-processing
  inequality), \texttt{Pythia.InfoTheory.BinaryEntropy} (Shannon
  binary entropy).
\item \textbf{Stochastic calculus:}
  \texttt{Pythia.Stochastic.ItoIsometryFiniteDim} (finite-dim Itô
  isometry), \texttt{Pythia.BDG} (Burkholder-Davis-Gundy at
  $p = 2$), \texttt{Pythia.MeasureTheory.}
  \texttt{ConditionalJensen}, \texttt{Pythia.StochasticApproximation.}
  \texttt{RobbinsSiegmund} / \texttt{RobbinsMonro} /
  \texttt{Dvoretzky}, \texttt{Pythia.TimeSeries.WoldDecomposition}
  \citep{wold1938}, \texttt{Pythia.TimeSeries.NeweyWest}.
\item \textbf{Quantum information:}
  \texttt{Pythia.Quantum.VonNeumannEntropyNonnegTwoState}
  (qubit von Neumann entropy non-negativity).
\item \textbf{Optimal transport:}
  \texttt{Pythia.OptimalTransport.WassersteinDistanceNonneg}
  (discrete L1 Wasserstein non-negativity).
\item \textbf{Game theory:}
  \texttt{Pythia.GameTheory.MinimaxTwoStrategyBound}
  (von Neumann two-strategy minimax inequality).
\item \textbf{Biology:} \texttt{Pythia.Bio.Population}
  (Hardy-Weinberg, Lotka-Volterra equilibria, SIR conservation),
  \texttt{Pythia.Bio.MichaelisMentenSaturation} (enzyme
  saturation), \texttt{Pythia.Bio.MassAction} (CRN ODE
  conservation), \texttt{Pythia.Bio.Phylogenetics}.
\item \textbf{Chemistry:} \texttt{Pythia.Chemistry.Arrhenius}
  (rate-law positivity), \texttt{Pythia.Chemistry.}
  \texttt{HendersonHasselbalch} (pH monotonicity),
  \texttt{Pythia.Chemistry.MassActionConservation}.
\item \textbf{Economics:} \texttt{Pythia.Economics.CobbDouglas}
  (constant returns to scale), \texttt{Pythia.Economics.CRRA}
  (utility concavity), \texttt{Pythia.Economics.CAPM}
  (linear pricing), \texttt{Pythia.Economics.RiskNeutralCall}
  (non-negativity), \texttt{Pythia.Economics.Walras}
  (excess-demand identity).
\item \textbf{Engineering:}
  \texttt{Pythia.Engineering.RCTimeConstant},
  \texttt{Pythia.Engineering.SignalEnergy},
  \texttt{Pythia.Engineering.PowerDissipation}.
\item \textbf{Mechanical:}
  \texttt{Pythia.Mechanical.HookeSpring}
  (spring-potential non-negativity).
\item \textbf{Control theory:}
  \texttt{Pythia.Control.Lyapunov} (scalar),
  \texttt{Pythia.Control.LyapunovDiscrete} (discrete-time),
  \texttt{Pythia.Control.LyapunovODE}.
\item \textbf{Operations research / queueing:}
  \texttt{Pythia.Queueing.ErlangB},
  \texttt{Pythia.Queueing.LittlesLaw},
  \texttt{Pythia.OR.LittlesLaw}.
\item \textbf{Numerical analysis:}
  \texttt{Pythia.Numerical.NewtonQuadraticIterPos},
  \texttt{Pythia.Numerical.PicardLindelof},
  \texttt{Pythia.Numerical.Lyapunov},
  \texttt{Pythia.Numerical.KKT},
  \texttt{Pythia.Numerical.Kahan}.
\item \textbf{Thermodynamics:}
  \texttt{Pythia.Thermodynamics.CarnotEfficiencyUpperBound}.
\item \textbf{Mathlib retags + actuarial:}
  \texttt{Pythia.MathlibTags} (AM-GM, Markov, Cauchy-Schwarz with
  empirical companions), \texttt{Pythia.Actuarial.\{Pareto,
  Weibull, LogNormal\}}.
\end{itemize}

\subsection{Five highlights}
\label{sec:library-highlights}

We name five representative results that exercise the tactic surface.
Module-level statements appear in
Appendix~\ref{sec:appendix-inventory}.

\begin{enumerate}[leftmargin=1.3em,itemsep=0.2em]
\item \textbf{Howard-Ramdas Ville bound.} The maximal inequality for
  the self-normalized confidence sequence
  \citep{howard2021time}, in
  \texttt{Pythia.HowardRamdasCS}. Closes via \texttt{anytime\_valid}.
\item \textbf{Betting-CS coverage.} The supermartingale coverage
  certificate of \citet{waudbysmith2024estimating}, in
  \texttt{Pythia.BettingCS}. Axiom-clean closure.
\item \textbf{Bernstein-Bennett MGF bridge.} The bounded-RV MGF bound
  in \texttt{Pythia.MGFBoundedSubGamma}, routing between the
  variance-aware sub-gamma tail and the Bennett MGF form.
\item \textbf{combine\_eprocesses\_avg.} Average-merge of e-processes
  \citep{ramdasgrunwaldvovkshafer2023, shinramdasrinaldo2024} in
  \texttt{Pythia.EDetector}. Axiom-clean closure with full
  reconstruction.
\item \textbf{Bretagnolle-Huber binary.} The binary-case
  total-variation lower bound in
  \texttt{Pythia.InfoTheory.} \texttt{BretagnolleHuberBinary}. Standard in
  information-theoretic lower-bound arguments.
\end{enumerate}

\paragraph{Pinning.}
The library is pinned to Lean 4.28 + Mathlib v4.28.0. The pin is
necessary: the modules consume Mathlib's measure theory and
probability subdirectories, which evolve quickly. We track Mathlib
release tags rather than \texttt{master} for reproducibility.
